FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-20 NewVersion/sections/Overview.tex
Overview

Notably, we track the static types by adding upcasts during reduction. In contrast to FJ,
where \lstinline|new C()| is a value, in \MIM{} such expression is not 
a value. Instead an expression of the form \lstinline|new C()| is reduced to
\lstinline|(C) new C()|, which is a value in \MIM{} and where the type
in the cast denotes the static type of the expression. This rule is
applied in the first reduction step. 
In the second reduction step, when 
\lstinline|shuffleAndDraw| is dispatched, the receiver
\lstinline|(DrawableSafeDeck) new DrawableSafeDeck()|
replaces the special variable \lstinline|this| by 
\lstinline|(Deck) new DrawableSafeDeck()|. Here, the static type used
in the cast denotes the origin of the \lstinline|shuffleAndDraw|
method (discovered during method lookup) used
for method dispatching, and the static type is changed to that. 
%The static type information about the origin
%of the method is helpful for future disambiguation. 
Later, in the fourth step,
\lstinline|((Deck) new DrawableSafeDeck()).draw()|
is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| above \lstinline|DrawableSafeDeck| and along path \lstinline|Deck|''. The meaning of ``above \lstinline|DrawableSafeDeck|'' implies its super types, and ``along path
\lstinline|Deck|'' specifies the branch. Finally, in the last
reduction step, we find the most
specific version of \lstinline|draw| in \lstinline|SafeDeck|. In this
sequence of
reduction steps, the cast that tracks the origin of
\lstinline|shuffleAndDraw| is crucial to unambiguously find the
correct implementation of \lstinline|draw|.
The formal procedure will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.
%Note that in C++, if we put the \lstinline[language=c++]|virtual| keyword in front of every \lstinline|draw| method, that means
%we enforce the dynamic dispatch, and it does the disambiguation correctly.
%Unfortunately, the C++ language, or rather the feature on triangle inheritance, has not been well formalized.
%In our language model, there are two types of casts: \emph{implicit casts} and \emph{explicit upcasts}. The implicit casts appear
%when passing arguments and returning expressions, they are also
%upcasts. Hence all casts are safe. \bruno{This discussion seems
%  inconplete. Either complete it or remove it.}

>>>>

Notably, we track the static types by adding upcasts during reduction. In contrast to FJ,
where \lstinline|new C()| is a value, in \MIM{} such expression is not 
a value. Instead an expression of the form \lstinline|new C()| is reduced to
\lstinline|(C) new C()|, which is a value in \MIM{} and where the type
in the cast denotes the static type of the expression. This rule is
applied in the first reduction step. 
In the second reduction step, when 
\lstinline|shuffleAndDraw| is dispatched, the receiver
\lstinline|(DrawableSafeDeck) new DrawableSafeDeck()|
replaces the special variable \lstinline|this| by 
\lstinline|(Deck) new DrawableSafeDeck()|. Here, the static type used
in the cast (\lstinline|Deck|) denotes the origin of the \lstinline|shuffleAndDraw|
method, which is discovered during method lookup. 
%The static type information about the origin
%of the method is helpful for future disambiguation. 
Later, in the fourth step,
\lstinline|((Deck) new DrawableSafeDeck()).draw()|
is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| above \lstinline|DrawableSafeDeck| and along path \lstinline|Deck|''. The meaning of ``above \lstinline|DrawableSafeDeck|'' implies its super types, and ``along path
\lstinline|Deck|'' specifies the branch. Finally, in the last
reduction step, we find the most
specific version of \lstinline|draw| in \lstinline|SafeDeck|. In this
sequence of
reduction steps, the cast that tracks the origin of
\lstinline|shuffleAndDraw| is crucial to unambiguously find the
correct implementation of \lstinline|draw|.
The formal procedure will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.
%Note that in C++, if we put the \lstinline[language=c++]|virtual| keyword in front of every \lstinline|draw| method, that means
%we enforce the dynamic dispatch, and it does the disambiguation correctly.
%Unfortunately, the C++ language, or rather the feature on triangle inheritance, has not been well formalized.
%In our language model, there are two types of casts: \emph{implicit casts} and \emph{explicit upcasts}. The implicit casts appear
%when passing arguments and returning expressions, they are also
%upcasts. Hence all casts are safe. \bruno{This discussion seems
%  inconplete. Either complete it or remove it.}
